---
title: "Final Project"
author: "Calvin Skalla"
date: "4/22/2020"
output: html_document
---

# Explaining the data
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(caret)
library(kableExtra)
library(cowplot)
library(ggthemes)
library(ggridges)

#Reading in the Data
white <- read.csv("data/winequality-white.csv", sep = ";")
red <- read.csv("data/winequality-red.csv", sep = ";")

#Combining and Clarifying Type
wine <- rbind(white, red) %>% tibble::rowid_to_column("ID") %>% mutate(Type = ifelse(ID <= 4898, 1, 0)) %>% select(-ID)

#Making quality a factor for easier visualization
wine$quality <- as.factor(wine$quality)

#Finding Outliers and Removing them

#normalizing data
preproc2 <- preProcess(wine, method = c("range"))
wine_minmax <- predict(preproc2, wine)

#Subsetting dataset for easier visualizations
twentypercent <- sample_frac(wine_minmax, size = 0.20)
```

Variable               Description
--------               -----------
`Sample Number`        Number given to each invdividual wine sample
`Quality`              Rating score on the quality of wine from 1-10
`Fixed Acidity`        Amount of acid from the grapes in grams per liter
`Volatile Acidity`     Amount of distillable acid in the wine in grams per liter
`Citric Acid`          Amount of citric acid present to add flavor in grams per liter
`Residual Sugar`       Leftover sugar from grapes after alcohol fermentation in grams per liter
`Chlorides`            Amount of salt in the wine in grams per liter
`Free Sulfur Dioxide`  Amount of $SO_2$ not bound to other molecules in milligrams per liter
`Total Sulfur Dioxide` Portion of free $SO_2$ plus the portion bounded to the other chemicals in milligrams per liter
`Density`              Mass of dissolved solids per volume in grams/$cm^3$ 
`pH`                   Measure of relative acidity vs. relative alkalinity of any liquid on a scale from 0-14
`sulfites`             Amount of preservative ($SO_2$) for antioxidant and antibacterial properties in grams per liter
`Alcohol`              Percent of alcohol content in the wine sample
--------               ------------
***

 <details>
  <summary> Click here to expand the correlation of each variable with Quality </summary>
```{r, echo=FALSE, message=FALSE}
#Correlation of Variables Table
cor_seq <- cor(x = wine_minmax[,1:11], y = as.numeric(wine_minmax$quality))
colnames(cor_seq) = c("Correlation")
variables = c("Fixed Acidity", "Volatile Acidity", "Citric Acid", "Residual Sugar", "Chlorides", "Free Sulfur Dioxide", "Total Sulfur Dioxide", "Density", "pH", "Sulfites", "Alcohol")

levels_of_significance = c("NSF", "(- -)", "(+)", "NSF", "(- -)", "NSF", "NSF", "(- - -)", "NSF", "NSF", "(+++)")

cor_data_frame <- data.frame("Variables" = variables, "Correlation" = cor_seq, "Significance" = levels_of_significance)
cor_data_frame <- cor_data_frame[order(-cor_data_frame$Correlation),]

cor_kbl <- kable(cor_data_frame, align = "lcc", caption = "Correlation of each variable with Quality", row.names = F) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive", fixed_thead = T), full_width = F) %>% row_spec(11, hline_after = T) %>% footnote(number = c("(+) refers to having a positive correlation with quality", "NSF refers to not having a sufficient correlation with quality", "(-) refers to having negative correlation with quality"))
cor_kbl
```
</details>

# Scatterplots to Demonstrate the Correlations of the Five Strongest Predictors
## Cannot figure out why geom_line() wont graph
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Showing alcohol levels against quality
alcohol_summary <- twentypercent %>% group_by(quality) %>% summarise(alcohol = mean(alcohol))

alcohol_scp <- ggplot(twentypercent, aes(x = quality, y = alcohol)) + geom_point(alpha = 1/10, position = position_jitter(height = 0.2, width = 0.2), color = "blue3") + geom_line(data = alcohol_summary, aes(quality, alcohol), size = 2) + geom_point(data = alcohol_summary, size = 2, color = "yellow")

#Showing density levels against quality
density_summary <- twentypercent %>% group_by(quality) %>% summarise(density = mean(density))

density_scp <- ggplot(twentypercent, aes(x = quality, y = density)) + geom_point(alpha = 1/10, position = position_jitter(height = 0.2, width = 0.2), color = "blue3") + geom_line(data = density_summary, size = 2, color = "yellow") + geom_point(data = density_summary, size = 2, color = "yellow")

#Showing volatile.acidity against quality
volatile_acidity_summary <- twentypercent %>% group_by(quality) %>% summarise(volatile.acidity = mean(volatile.acidity))

volatile_acidity_scp <- ggplot(twentypercent, aes(x = quality, y = volatile.acidity)) + geom_point(alpha = 1/10, position = position_jitter(height = 0.2, width = 0.2), color = "blue3") + geom_line(data = volatile_acidity_summary) + geom_point(data = volatile_acidity_summary, size = 2, color = "yellow")

#Showing chlorides against quality
chloride_summary <- twentypercent %>% group_by(quality) %>% summarise(chlorides = mean(chlorides))

chlorides_scp <- ggplot(twentypercent, aes(x = quality, y = chlorides)) + geom_point(alpha = 1/10, position = position_jitter(height = 0.2, width = 0.2), color = "blue3") + geom_line(data = chloride_summary, size = 2, color = "yellow") + geom_point(data = chloride_summary, size = 2, color = "yellow")

#Showing citric acid against quality
citric_acid_summary <- twentypercent %>% group_by(quality) %>% summarise(citric.acid = mean(citric.acid))

citric_acid_scp <- ggplot(twentypercent, aes(x = quality, y = citric.acid)) + geom_point(alpha = 1/10, position = position_jitter(height = 0.2, width = 0.2), color = "blue3") + geom_line(data = citric_acid_summary, size = 2) + geom_point(data = citric_acid_summary, size = 2, color = "yellow")

#Plotting all 5 plots nicely on a grid
plot_grid(alcohol_scp, density_scp, volatile_acidity_scp, chlorides_scp, citric_acid_scp, ncol = 2, nrow = 3)
```

# Use a Quality Cut to show differences in more varying qualities
```{r, echo=FALSE, message=FALSE, warning=FALSE}

```

# Use Violin Plots or Density Ridges to show distributions over quality for Non-significant variables
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Density Ridge plots for the three negative non-significant predictors
fixed_acidity_dens <- ggplot(wine_minmax) + geom_density_ridges(aes(x = fixed.acidity, y = quality))

total_sulfur_dio_dens <- ggplot(wine_minmax) + geom_density_ridges(aes(x = total.sulfur.dioxide, y = quality))

residual_sugar_dens <- ggplot(wine_minmax) + geom_density_ridges(aes(x = residual.sugar, y = quality))

plot_grid(fixed_acidity_dens, total_sulfur_dio_dens, residual_sugar_dens, ncol = 2, nrow = 2)

# Violin plots for the three positive non-significant predictors
free_sulfur_dio_vio <- ggplot(twentypercent, aes(x = quality, y = free.sulfur.dioxide)) + geom_violin() + coord_flip()

sulphates_vio <- ggplot(twentypercent, aes(x = quality, y = sulphates)) + geom_violin() + coord_flip()

pH_vio <- ggplot(twentypercent, aes(x = quality, y = pH)) + geom_violin() + coord_flip()

plot_grid(pH_vio, sulphates_vio, free_sulfur_dio_vio, ncol = 2, nrow = 2)
```

