---
title: "Predicting Red or White Wines"
author: "Calvin Skalla"
date: "4/15/2020"
output: html_document
---

# Data Cleaning
## 1 refers to White wine while 0 refers to Red Wine
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(knitr)
library(ggthemes)
library(cowplot)
library(plyr)
library(readr)
library(GGally)
library(mlbench)
library(caret)
library(glmnet)

#Reading in the Wine
white_wine <- read.csv("data/winequality-white.csv", sep = ";")
red_wine <- read.csv("data/winequality-red.csv", sep = ";")

#Combining and clarifying the red vs white
wine_data <- rbind(white_wine, red_wine) %>% tibble::rowid_to_column("ID") %>% mutate(Type = ifelse(ID <= 4898, 1, 0)) %>% select(-ID)

#Eliminating Outliers
#Fixed Acidity Outliers
#fa_outliers <- boxplot(wine_data$fixed.acidity)$out
#wine_data <- wine_data[-which(wine_data$fixed.acidity %in% fa_outliers),]
#Volatile Acidity Outliers
#va_outliers <- boxplot(wine_data$volatile.acidity)$out
#wine_data <- wine_data[-which(wine_data$volatile.acidity %in% va_outliers),]
#Citric Acid Outliers
#ca_outliers <- boxplot(wine_data$citric.acid)$out
#wine_data <- wine_data[-which(wine_data$citric.acid %in% ca_outliers),]
#Residual Sugar Outliers
#rs_outliers <- boxplot(wine_data$residual.sugar)$out
#wine_data <- wine_data[-which(wine_data$residual.sugar %in% rs_outliers),]
#Free SUlfur Dioxide Outliers
#fsd_outliers <- boxplot(wine_data$free.sulfur.dioxide)$out
#wine_data <- wine_data[-which(wine_data$free.sulfur.dioxide %in% fsd_outliers),]
#Total SUlfur Dioxide Outliers
#tsd_outliers <- boxplot(wine_data$total.sulfur.dioxide)$out
#wine_data <- wine_data[-which(wine_data$total.sulfur.dioxide %in% tsd_outliers),]
#Density Outliers
#d_outliers <- boxplot(wine_data$density)$out
#wine_data <- wine_data[-which(wine_data$density %in% d_outliers),]
#pH Outliers
#pH_outliers <- boxplot(wine_data$pH)$out
#wine_data <- wine_data[-which(wine_data$pH %in% pH_outliers),]
#Sulphates Outliers
#s_outliers <- boxplot(wine_data$sulphates)$out
#wine_data <- wine_data[-which(wine_data$sulphates %in% s_outliers),]
#Alcohol Outliers
#a_outliers <- boxplot(wine_data$alcohol)$out
#wine_data <- wine_data[-which(wine_data$alcohol %in% a_outliers),]

#normalizing data
preproc2 <- preProcess(wine_data, method = c("range"))
w_data_minmax <- predict(preproc2, wine_data)

#Splitting data into train/test
train_index <- sample(1:nrow(w_data_minmax), 0.8 * nrow(w_data_minmax))
test_index <- setdiff(1:nrow(w_data_minmax), train_index)
train <- w_data_minmax[train_index, ]
test <- w_data_minmax[test_index, ]
#count(train$Type)
#count(test$Type)

```

# First Model
### Type ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol 
```{r, echo=FALSE, warning=FALSE, message=FALSE}
#All available variables
first_mod = glm(Type ~ fixed.acidity + volatile.acidity + citric.acid
             + residual.sugar + chlorides + free.sulfur.dioxide 
             + total.sulfur.dioxide + density + pH + sulphates
             + alcohol, data = train, family = binomial)

#summary(first_mod)

first_probs = predict(first_mod, test, type = "response")
first_pred = rep(0, 1300)
first_pred[first_probs>0.5] = 1
table(first_pred, test$Type)
mean(first_pred==test$Type)
```

# Second Model
### Type ~ volatile.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + alcohol
```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Only significant variables
second_mod = glm(Type ~ volatile.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + alcohol, data = train, family = binomial)

#summary(second_mod)
second_probs = predict(second_mod, test, type = "response")
second_pred = rep(0, 1300)
second_pred[second_probs>0.5] = 1
table(second_pred, test$Type)
mean(second_pred==test$Type)
```

# Singular Models
### Type ~ total.sulfur.dioxide | Type ~ chlorides | Type ~ volatile.acidity
```{r, echo=FALSE, warning=FALSE, message=FALSE}
sing1_mod = glm(Type ~ total.sulfur.dioxide, data = train, family = binomial)

sing1_probs = predict(sing1_mod, test, type = "response")
sing1_pred = rep(0, 1300)
sing1_pred[sing1_probs>0.5] = 1
table(sing1_pred, test$Type)
mean(sing1_pred==test$Type)

sing2_mod = glm(Type ~ chlorides, data = train, family = binomial)

sing2_probs = predict(sing2_mod, test, type = "response")
sing2_pred = rep(0, 1300)
sing2_pred[sing2_probs>0.5] = 1
table(sing2_pred, test$Type)
mean(sing2_pred==test$Type)

sing3_mod = glm(Type ~ volatile.acidity, data = train, family = binomial)

sing3_probs = predict(sing3_mod, test, type = "response")
sing3_pred = rep(0, 1300)
sing3_pred[sing3_probs>0.5] = 1
table(sing3_pred, test$Type)
mean(sing3_pred==test$Type)

```

# Visualization of Singular Models
```{r, echo=FALSE, warning=FALSE, message=FALSE}
tsd_lr <- ggplot(train, aes(x = total.sulfur.dioxide, y = Type, color = alcohol)) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  ggtitle("Logistic regression model fit") +
  xlab("Total Sulfur Dioxide Levels") +
  ylab("Type of Wine")

c_lr <- ggplot(train, aes(x = chlorides, y = Type, color = alcohol)) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  ggtitle("Logistic regression model fit") +
  xlab("Chloride Levels") +
  ylab("Type of Wine")

va_lr <- ggplot(train, aes(x = volatile.acidity, y = Type, color = alcohol)) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  ggtitle("Logistic regression model fit") +
  xlab("Volatile Acidity Levels") +
  ylab("Type of Wine")

plot_grid(tsd_lr, c_lr, va_lr, ncol = 2, nrow = 2)
```
